Verification of Stencil Computations through Spatial Specifications

### 2016-07-07 paper.tex
section 4 edits

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
%\usepackage{subcaption}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

@@

\begin{figure}[t]
\vspace{-0.1em}
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{N}_{>0}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v \\
\nonterm{regionDec} ::= &  \texttt{region} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\[0.4em]
%\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}
%\\[1em]
\nonterm{spec} ::= & [\nonterm{approx},] \; [\nonterm{mult},] \; \nonterm{region} \\
\nonterm{mult} ::= & \term{readOnce} \\
\nonterm{approx} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \nonterm{rvar} \; \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenRefl{\dimTy}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \;\stenFwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenBwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenCen{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{region} \; \term{+} \; \nonterm{region}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{region} \; \term{*} \; \nonterm{region}} \\[0.5em]
%\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{rvar}}  
%\\[0.5em]
%\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
%  \\[0.5em]
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+\\[-1em]
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\vspace{-0.5em}
\end{figure}

@@

\begin{figure}
\vspace{-0.2em}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\textit{region} \equiv \textit{region}'$} \\[0.7em]
\begin{array}{rll}
(\textsc{F\, \texttt{+} \,F}) \;\; &
\stenFwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,B}) \;\; &
\stenBwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenBwdSR{n}{d}{r} \; \texttt{+} \; \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,C}) \;\; &
\stenCenSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
\equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \; \stenCenSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,F}) \;\; & \stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,B}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,F}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenBwdSR{n}{d}{s}
  \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,F}) \;\; &
\stenFwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenFwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,B}) \;\; &
\stenBwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenBwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,C}) \;\; &
\stenCenS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenCenSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{irrefl}) \;\; & (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \;
\equiv \; R \, \term{*} \, S \\[0.4em]
%
%
(\texttt{+}\textsc{idem}) \;\; & S \; \texttt{+} \; S \; \equiv \; S \\[0.4em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.4em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.4em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.4em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.4em]
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)
\end{array}
%
%\framebox{${\textit{specDec} :: \texttt{v}} \equiv
%{\textit{specDec'} :: \texttt{v}}$} \\[0.7em]
%\begin{array}{rl}
% MUTUAL
%(\textsc{mutual}) \; &
%\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
%  :: \texttt{w}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
%  \texttt{v} \\[0.5em]
%(\textsc{coalesce}) \; &
%\texttt{stencil} \; \texttt{dependency} (\bar{v})
%  :: \texttt{v} \\
%& ; \texttt{stencil} \; \texttt{dependency} (\bar{w})
%  :: \texttt{v}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
%  \texttt{v} \\[0.5em]
%\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications (regions)}
\vspace{-1em}
\label{fig:equations}
\end{figure}

@@

Let $U$ range over the finite maps generated by \textsf{analyse}.  For
each $x \in \mathsf{dom}(U)$, the next step of the algorithm
constructs a set of $n$-dimensional rectangles covering all contiguous groups of schemes
in $U(x)$.

@@

\begin{definition}[Spans]
  A \emph{span} represents an $n$-dimensional box (\emph{hyper-rectangle}) as
  a pair of $n$-dimensional vectors (or represented as a $2 \times n$
  matrix) with values drawn from $\domainVal{}$. This  
   gives the co-ordinates of a lower-bound vertex (first
  column) and the upper-bound vertex (second column). We use the
  notation $\vect{x} = [\vect{x}^L \; \vect{x}^U]$, where $\vect{x}^L$
  and $\vect{x}^U$ give the lower and upper bounds of the span.
In the rest, $\vect{x}, \vect{y}$ range over spans. 
For a span $\vect{x}$, we write $\vect{x}_{i:j}$ for the sub-span
comprises the subvectors $[\vect{x}^L_{i:j} \; \vect{x}^U_{i:j}]$ from entry $i$ to
entry $j$ (inclusive).
\end{definition}

\noindent
The indexing schemes of $U(x)$ are converted to \emph{unit} spans $u \mapsto [u \; 
u]$, defining the lower and upper bound.
 For the five-point stencil example there are five $1 \times 1$ spans.
Figure~\ref{fig:inference-steps-informal}(a) illustrates these.

\begin{figure}[t]
\vspace{-0.3em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0em]
\begin{tikzpicture}[scale=0.8]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\footnotesize{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} \\[0em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture} \\[-0.5em] (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture} \\[-0.5em] (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in
specification inference}
\label{fig:inference-steps-informal}
\vspace{-0.75em}
\end{figure}

% taking permustations is a systematic way of producing all rotations
% of the n-dimensional space - forward ref to fold
The next step of the algorithm repeatedly coalesces any contiguous spans
until a fixed point is reached, resulting in a set of spans covering
over the original space of schemes. Two $n$-dimensional spans $\vect{x}$ and $\vect{y}$
are contiguous if for all but one $j \in \{1, \ldots, n\}$ then
$\vect{x}_j = \vect{y}_j$ and for one $j$ then $\vect{x}^U_1 + 1 =
\vect{y}^U$. For example, unit spans
$\left[\vtwoh{-1}{0}\; \vtwoh{-1}{0}\right]$ and $[\vtwoh{0}{0} \; \vtwoh{0}{0}]$ 
are contiguous in the first dimension and can be coalesced
into the span $[\vtwoh{-1}{0} \; \vtwoh{0}{0}]$.

\begin{definition}\label{def:span-coalesc} Our operator for span
  coalescing $\bullet$ is defined for spans contiguous in the first dimension:
%%
\begin{align*}
\vect{x} \bullet \vect{y}
= &
\begin{cases}
(\vect{x}^L, \vect{y}^U) & \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
\vect{x}_{2:n} = \vect{y}_{2:n}) \\
\bot  & \textit{otherwise}
\end{cases}
% $m$-dimensions into coalesced spans which vary in
% $m+1$-dimensions.
\end{align*}
%%
\end{definition}
In order to coalesce regions which are contiguous in other dimensions
other than the first, we generate all rotations of the spans in
$n$-dimensional space by permutation, and fold together contiguous
regions using the above $\bullet$ operation. For our example,
from the unit spans in Figure~\ref{fig:inference-steps-informal}(a)
the contiguous spans in Figure~\ref{fig:inference-steps-informal}(b)
are generated. Then any spans that are contained with any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another) (Figure~\ref{fig:inference-steps-informal}(c).

This procedure is iterated until a fixed point is reached. In this
example this is reached in the first step. The final spans
are used to synthesis concrete stencil specification syntax, where
each span is product \texttt{*} of regions and multiple spans are
combined with the region sum \texttt{+}.

More formally, the algorithm to create contiguous spans, covering the space of
indices, now proceeds as follows.
Firstly, for all $v \in \mathsf{dom}(M)$ a new map $M'$ is
created where each vector is mapped to the trivial \emph{unit span} is
created by pairing a vector with itself:
%
\begin{align*}
N(v) = \{[\vect{x} \;
  \vect{x}] \mid \vect{x} \in M(v)\}
\end{align*}
%
For our example, this yields:
%%
\begin{align*}
N(\texttt{a}) = \stwo{0}{0}{0}{0} \stwo{1}{0}{1}{0} \stwo{-1}{0}{-1}{0} \stwo{0}{1}{0}{1} \stwo{0}{-1}{0}{-1}
\end{align*}
%%
A fixed point is then computed for the following
procedure \spanOp which, per variable
in the map, coalesces spans into contiguous regions in
the $n$-dimensional space. That is, we compute
 $(\mu \; \spanOp) N$ where \spanOp
is defined as follows for each variable in the map:
%
\begin{enumerate}

\item Compute all permutations on the column vectors in a span, \ie{},
  $[\vect{x}^L \vect{x}^U] \mapsto [\pi\vect{x}^L \, \pi\vect{x}^U]$
for a permutation $\pi$. For each permutation function $\pi^n_i$
(the $i$-th permutation for vectors of size $n$) we pair the
permutation function with the set of permuted spans so that
the spans can be un-permuted later.
%
\[
P(v) = \bigcup_{i \in n!} (\pi^n_{i} , \; \{[\pi^n_i
\vect{x}^L \; \pi^n_i\vect{x}^U] \, \mid \, [\vect{x}^L \; \vect{x}^U]
\leftarrow N(v)\}
\]
%
For our example, this yields:
%
\begin{align*}
P(\texttt{a}) =
\{&(\pi^2_1, \{ \stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{-1}{0}{-1}{0},
\stwo{0}{1}{0}{1},
\stwo{0}{-1}{0}{-1} \})
\\
&(\pi^2_2, \{
 \stwo{0}{0}{0}{0},
 \stwo{0}{1}{0}{1},
 \stwo{0}{-1}{0}{-1},
 \stwo{1}{0}{1}{0},
 \stwo{-1}{0}{-1}{0}\})\}
\end{align*}
%
where $\pi^2_1$ is the identity permutation and $\pi^2_2$ is the
permutation flips the order of the two elements in the column
vectors.

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
\usepackage{enumitem}
%\usepackage{subcaption}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

@@

\begin{figure}[t]
\vspace{-0.1em}
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{N}_{>0}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v \\
\nonterm{regionDec} ::= &  \texttt{region} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\[0.4em]
%\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}
%\\[1em]
\nonterm{spec} ::= & [\nonterm{approx},] \; [\nonterm{mult},] \; \nonterm{region} \\
\nonterm{mult} ::= & \term{readOnce} \\
\nonterm{approx} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \nonterm{rvar} \; \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenRefl{\dimTy}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \;\stenFwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenBwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenCen{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{irreflexive}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{region} \; \term{+} \; \nonterm{region}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{region} \; \term{*} \; \nonterm{region}} \\[0.5em]
%\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{rvar}}  
%\\[0.5em]
%\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
%  \\[0.5em]
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+\\[-0.5em]
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\vspace{-0.8em}
\end{figure}

@@

\begin{figure}
%\vspace{-0.2em}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
%\framebox{$\textit{region} \equiv \textit{region}'$} \\[0.7em]
\begin{array}{rll}
(\textsc{F\, \texttt{+} \,F}) \;\; &
\stenFwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,B}) \;\; &
\stenBwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenBwdSR{n}{d}{r} \; \texttt{+} \; \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,C}) \;\; &
\stenCenSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
\equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \; \stenCenSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,F}) \;\; & \stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,B}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,F}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenBwdSR{n}{d}{s}
  \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,F}) \;\; &
\stenFwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenFwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,B}) \;\; &
\stenBwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenBwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,C}) \;\; &
\stenCenS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenCenSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{irrefl}) \;\; & (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \;
\equiv \; R \, \term{*} \, S \\[0.4em]
%
%
(\texttt{+}\textsc{idem}) \;\; & S \; \texttt{+} \; S \; \equiv \; S \\[0.4em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.4em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.4em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.4em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.4em]
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)
\end{array}
%
%\framebox{${\textit{specDec} :: \texttt{v}} \equiv
%{\textit{specDec'} :: \texttt{v}}$} \\[0.7em]
%\begin{array}{rl}
% MUTUAL
%(\textsc{mutual}) \; &
%\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
%  :: \texttt{w}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
%  \texttt{v} \\[0.5em]
%(\textsc{coalesce}) \; &
%\texttt{stencil} \; \texttt{dependency} (\bar{v})
%  :: \texttt{v} \\
%& ; \texttt{stencil} \; \texttt{dependency} (\bar{w})
%  :: \texttt{v}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
%  \texttt{v} \\[0.5em]
%\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications (regions)}
\vspace{-1em}
\label{fig:equations}
\end{figure}

@@

Let $U$ range over the finite maps computed by \textsf{analyse}.  For
each $x \in \mathsf{dom}(U)$, the algorithm 
constructs next a set of $n$-dimensional rectangles covering all contiguous groups of schemes
in $U(x)$.

@@

\begin{definition}[Spans]
  A \emph{span} represents an $n$-dimensional box
  (\emph{hyper-rectangle}) as a pair of $n$-dimensional vectors (or
  represented as a $2 \times n$ matrix) with values drawn from
  $\domainVal{}$. This gives the co-ordinates of a lower-bound vertex
  (first comonent) and the upper-bound (second component). We write 
  spans as $\vect{x} = [\vect{x}^L \; \vect{x}^U]$, where $\vect{x}^L$
  and $\vect{x}^U$ give the lower and upper bounds.  In
  the rest, $\vect{x}, \vect{y}$ range over spans.  For a span
  $\vect{x}$, we write $\vect{x}_{i:j}$ for the sub-span comprises the
  subvectors $[\vect{x}^L_{i:j} \; \vect{x}^U_{i:j}]$ from entry $i$
  to entry $j$ (inclusive).
\end{definition}

\noindent
Each indexing scheme in $U(x)$ is converted to a \emph{unit} span $u \mapsto [u \; 
u]$, defining the lower and upper bound.
 For the five-point stencil example there are five $1 \times 1$ spans.
Figure~\ref{fig:inference-steps-informal}(a) illustrates these.

\begin{figure}[t]
\vspace{-0.3em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0em]
\begin{tikzpicture}[scale=0.8]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\footnotesize{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} \\[0em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture} \\[-0.5em] (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture} \\[-0.5em] (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in
specification inference}
\label{fig:inference-steps-informal}
\vspace{-0.8em}
\end{figure}

% taking permustations is a systematic way of producing all rotations
% of the n-dimensional space - forward ref to fold
The next step repeatedly coalesces any contiguous spans
until a fixed point is reached, resulting in a set of (possibly
overlapping) spans covering the original space of schemes. Two $n$-dimensional spans $\vect{x}$ and $\vect{y}$
are contiguous if for all but one $j \in \{1, \ldots, n\}$ then
$\vect{x}_j = \vect{y}_j$ and for one $j$ then $\vect{x}^U_1 + 1 =
\vect{y}^U$. For example, spans
$\left[\vtwoh{-1}{0}\; \vtwoh{-1}{0}\right]$ and $[\vtwoh{0}{0} \; \vtwoh{1}{0}]$ 
are contiguous in the first dimension and can be coalesced
into the span $[\vtwoh{-1}{0} \; \vtwoh{1}{0}]$.

\begin{definition}\label{def:span-coalesc} Span
  coalescing $\bullet$ is defined for spans contiguous in the first dimension:
%%
\begin{align*}
\vect{x} \bullet \vect{y}
= &
\begin{cases}
(\vect{x}^L, \vect{y}^U) & \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
\vect{x}_{2:n} = \vect{y}_{2:n}) \\
\bot  & \textit{otherwise}
\end{cases}
% $m$-dimensions into coalesced spans which vary in
% $m+1$-dimensions.
\end{align*}
%%
\end{definition}
\noindent
To coalesce regions which are contiguous in other dimensions
other than the first, we generate all rotations of the spans in
$n$-dimensional space by permutation on spans, and fold together contiguous
regions using $\bullet$ (above). This 
generates all possible ways of coalescing regions. For our example,
the unit spans in Figure~\ref{fig:inference-steps-informal}(a)
are colaesced into the contiguous spans in
Figure~\ref{fig:inference-steps-informal}(b).
This procedure is iterated until a fixed point is reached. In this
example this is reached in the first step. 
 Then any spans that are completely contained by any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another), Figure~\ref{fig:inference-steps-informal}(c).
The final spans are used to synthesis concrete stencil specification
syntax (Section~\ref{sec:inf-step4}), where
each span becomes a product \texttt{*} of regions which are 
combined by the sum \texttt{+}.

More formally the algorithm proceeds as follows.
For all $x \in \mathsf{dom}(U)$, let $M = U(x)$. Every scheme in $M$ 
is mapped to a unit span
by $\textsf{unitSpans}(M) = \{[u \; u] \mid u\!\in\!M\}$. For our example:
%%
\begin{align*}
N(U(\texttt{a})) = \stwo{0}{0}{0}{0} \stwo{1}{0}{1}{0} \stwo{-1}{0}{-1}{0} \stwo{0}{1}{0}{1} \stwo{0}{-1}{0}{-1}
\end{align*}
%%
In the following we let $M_0 = N(U(\texttt{a}))$ for our example.

A fixed point of the following function, \spanOp{}, is taken (per variable
in the map U) to coalesces spans into contiguous regions, \ie{}, we compute
 $(\mu \, \spanOp) \; N(M)$ where \spanOp{} is defined by: \\[-0.5em]

\noindent
\textit{1).\;} Compute all permutations on the column vectors in a span, \ie{},
  $[\vect{x}^L \vect{x}^U] \mapsto [\pi\vect{x}^L \, \pi\vect{x}^U]$
for a permutation $\pi$. For each permutation function $\pi^n_i$
(the $i$-th permutation for vectors of size $n$) we pair the
permutation function with the set of permuted spans so that
the spans can be un-permuted later.
%
\begin{equation*}
P(M) = \bigcup_{i \in n!} (\pi^n_{i} , \; \{[\pi^n_i
\vect{x}^L \; \pi^n_i\vect{x}^U] \, \mid \, [\vect{x}^L \; \vect{x}^U]
\leftarrow M\}
\end{equation*}
%
This corresponds to all rotations of the space.  For our example:
%
\begin{align*}
P(M_0) =
\{&(\pi^2_1, \{ \stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{-1}{0}{-1}{0},
\stwo{0}{1}{0}{1},
\stwo{0}{-1}{0}{-1} \})
\\
&(\pi^2_2, \{
 \stwo{0}{0}{0}{0},
 \stwo{0}{1}{0}{1},
 \stwo{0}{-1}{0}{-1},
 \stwo{1}{0}{1}{0},
 \stwo{-1}{0}{-1}{0}\})\}
\end{align*}
%
where $\pi^2_1$ is the identity permutation and $\pi^2_2$ is the
permutation flips the order of the two elements in the 
vectors. \\
